Bayesian exponential random graph modeling of KWL Burial network

### 2020-11-06 analysis/paper/paper.Rmd
add citations to Intro

<!--burial study- network analysis-->
Burials analysis has long developed in archaeology as an approach to study past social structure through mortuary practices, such as cultural and biological records in burial contexts, that can represent social status or identity of the deceased [@Binford1971; @Saxe1970]. Despite the criticism for overlooking the burial manipulation by the living that does not necessarily connect to the real life [@Hodder1980; @Pearson1982], with comparisons between different archaeological contexts of death and life in mind, burials can still provide valuable information to infer past societies [@Chapman2003]. For example, mortuary practices, including burial forms, grave goods, and ritual behaviors, that structure the material configuration of burials can represent social relations between members of a community, allowing the inference for social structures [@Byrd1995; @Seikel2011]. To study social relations, the recent development of network science and graph theory provides many new tools for the exploration and analyzing the structure of relational data [@Brandes2013]. This means we can answer anthropological questions about social differentiation, and social inequality or complexity based on our assumptions of archaeological data [@Ehrhardt2009; @Gamble2002; @Janes2013; @Johansen2015]. For example, burial goods can reflect social practices in broader cultural contexts to represent personal wealth, social roles, or traditional customs. The flow of material goods  Burials with the same goods may indicate some similar underlying social relations where individuals share similar access to trade, exchange, and gifting networks [eg. @Coward2013, pp. 252]. Identifying the structure of relationships among burials can give insights into the structure of social organization.

To measure relations and patterns underlying the relations, social network analysis (SNA), is a useful method that has been increasingly applied by archaeologists to study relationship between humans, or sites in past societies [@Mills2017]. A social network can be expressed as a graph consisting of a set of socially-relevant nodes, or actors, connected by edges representing one or more relations, for example, friendship, collaborations, information flow, trade ties, or any forms of connection of interest [@Marin2011; @Wasserman1994]. The basic unit of relationship is called a dyad, defined as a connection between two distinct actors [@Lee2016]. In archaeology, actors can be people, groups, objects, places, or events, with ties built on similarity, proximity, and co-presence of material culture based on archaeological theories for relationships [@Mills2017; @Brughmans2018]. This allows answering questions related to interaction and exchange, identities, and power relationships between social groups or archaeological sites by quantifying and visualizing social relations from archaeological data. There are two common approaches of archaeological applications, one is analyzing structures of networks built on archaeological data and finding patterns, and another is modeling theory to represent and test dependence of relationships [@Brughmans2018]. The modeling approach is useful to statistically test hypotheses about the relationships between actors to examine the structure of network and major factors that affect the formation of the structure [@Freeman2004; @Salvini2010]. Those approaches are widely used for different topics, such as exploring hunter-gatherer exchange network in the Kuril Islands based on ceramic compositions [@Erik2015]; long term inter-site relationship from the Epipalaeolithic to the early Neolithic in Near East according to trade items [@Coward2013]; hierarchical communication systems in the Kofun period in Japan through prestige goods [@Mizoguchi2013].

@@

# Exponential random graph models in a Bayseisan framework

@@

<!--Model fitting and prior setting--> 
After the specification, the network is simulated in a Baysesian framework using Markov chain Monte Carlo (MCMC) algorithm. MCMC algorithm is an estimation of posterior distribution of interest through directly random sampling the posterior without assuming the prior is a normal distribution. We can obtain a distribution by constructing a Markov chain that describes a sequence of moves from current state to the next state following probabilistic rules based on algorithms. This enables a random or stochastic simulation in a long run without depending on the previous move. More chains ensures a more desirable distribution close to the target distribution that we study, which is called the convergence of the Markov chains. In Bayesian ERGMs, MCMC first selects a set of edges (or a set of empty pair of actors) with equal probability, and than switch to a pair of actors at random within the chosen set [@Caimo2011]. In our case, we set the number of chains to six. For each chain, the number of burn-in iterations is 200 and the number of iterations after the burn-in is 2000. We set the number of iterations used to simulate a network y′ at each iteration to 10,000. For prior, the normal distribution is agreed as a suitable prior model in network analysis that assumes networks with low density and high transitivity commonly found in real world [@Caimo2017]. To

>>>>

<!--burial study- network analysis-->
Burials analysis has long developed in archaeology as an approach to study past social structure through mortuary practices, such as cultural and biological records in burial contexts, that can represent the social status or identity of the deceased [@Binford1971; @Saxe1970]. Despite the criticism for overlooking the burial manipulation by the living that does not necessarily connect to the real life [@Hodder1980; @Pearson1982], with comparisons between different archaeological contexts of death and life in mind, burials can still provide valuable information to infer past societies [@Chapman2003]. For example, mortuary practices, including burial forms, grave goods, and ritual behaviors, that structure the material configuration of burials can represent social relations between members of a community, allowing the inference for social structures [@Byrd1995; @Seikel2011]. To study the relations among burials, the recent development of network science provides many new tools to visualize and analyze relationships. Network science is the study of relational data where a phenomenon is conceptualized as a network through two steps, abstraction and representation, based on our specification in terms of network theories [@Brandes2013]. That is, a network can be viewed as a patterned aggregation that includes individual elements, pair-wise relationships (the dyads), and overall structure showing global patterning represented in data.

A network is generally visualized as a graph consisting of a set of socially-relevant nodes, or actors, connected by edges representing one or more relations, such as friendship, collaborations, information flow, trade ties, or any other forms of connection of interest [@Marin2011; @Wasserman1994]. In archaeology, actors can be people, groups, objects, places, or events, with ties built on similarity, proximity, and co-presence of material culture based on theories for relationships [@Mills2017; @Brughmans2018]. For example, past trade can be conceptualized as a network of individual entities connected by shared similarity, the flow of goods, to represent their interactions [@Collar2015, pp. 4]. Similar concepts can be applied to a wide range of archaeological data with relational assumptions, such as burial analysis. Burial goods, especially for high value goods, can reflect social practices in broader cultural contexts to represent personal wealth or social status that can infer social differentiation or complexity [@Gamble2002; @Janes2013]. Burials with the same prestige goods could indicate some underlying social relations where individuals share similar access to trade, exchange, and gifting networks according to their status [eg. @Coward2013, pp. 252]. This enables the exploration of the structure of the past social organization by identifying the relationships among burials.

Network analysis has been increasingly applied by archaeologists in recent years to deal with past interactions and explore the underlying mechanism. There are two common methodological approaches, one is directly examining network properties, such as node position and overall structures of networks built on archaeological data, to find meaningful patterns [@Mills2017]; another one is modeling theory to represent and test dependence of relationships [@Brughmans2018]. Node position is important to identify the role of a node in a network, such as its centrality, representing the individual influence or social prominence in a group, while network structure is helpful to generalize relationship patterns. Recent applications include long term inter-site relationship from the Epipalaeolithic to the early Neolithic in Near East according to trade items [@Coward2013], and hierarchical communication systems in the Kofun period in Japan through prestige goods [@Mizoguchi2013]. For modeling approaches, they provide a robust way to statistically test hypotheses about the relationships between actors to examine the structure of the network and major factors that affect the formation of the structure [@Freeman2004; @Salvini2010]. Examples include testing hunter-gatherer exchange network across the Kuril Islands based on ceramic compositions [@Erik2015], and modeling and comparing chronological and social hypotheses to explain the differences between burials [@Sosna2012]. This paper will use modeling approaches for the relationship between burials in northeastern Taiwan to explore the indirect impact of foreign contacts.

@@

# Exponential random graph models in a Baysesian framework

@@

<!--Model fitting and prior setting--> 
After the specification, the network is simulated in a Baysesian framework using Markov chain Monte Carlo (MCMC) algorithm. MCMC algorithm is an estimation of posterior distribution of interest through directly random sampling the posterior without assuming the prior is a normal distribution. We can obtain a distribution by constructing a Markov chain that describes a sequence of moves from current state to the next state following probabilistic rules based on algorithms. This enables a random or stochastic simulation in a long run without depending on the previous move. More chains ensures a more desirable distribution close to the target distribution that we study, which is called the convergence of the Markov chains. In Bayesian ERGMs, MCMC first selects a set of edges (or a set of empty pair of actors) with equal probability, and than switch to a pair of actors at random within the chosen set [@Caimo2011]. In our case, we set the number of chains to six. For each chain, the number of burn-in iterations is 200 and the number of iterations after the burn-in is 2000. We set the number of iterations used to simulate a network y′ at each iteration to 10,000. For prior, the normal distribution is agreed as a suitable prior model in network analysis that assumes networks with low density and high transitivity commonly found in real world [@Caimo2017].
